Can Cultural Transmission Explain The Evolution of Cooperative Behavior?

### 2020-04-18 ms/main.tex

Horizontal cultural transmission occurs between peers. 
It may occur between social partners with probability $\alpha$, or between a random pair with probability $1-\alpha$ (see~\textbf{Fig.~\ref{fig:horizontal}}).
Horizontal transmission is not always successful, as one peer may reject the other's phenotype. The probability for successful transmission of phenotypes $A$ and $B$ are $T_A$ and $T_B$, respectively.

Therefore, the frequency $p'$ of phenotype $A$ among adults in the next generation, after horizontal transmission, is 
\begin{equation}\label{eq:nextgen_adults}
\begin{aligned}
p'
& = \hat{p}^2 [\alpha + (1-\alpha)(\hat{p} + (1-\hat{p})(1-T_B))] \\
& + \hat{p}(1-\hat{p}) [\alpha(1-T_B) + (1-\alpha)(\hat{p} + (1-\hat{p})(1-T_B))] \\
& + (1-\hat{p})\hat{p} [\alpha T_A + (1-\alpha) \hat{p} T_A ] \\
& + (1-\hat{p})^2 [(1-\alpha) \hat{p} T_A]
\end{aligned}
\end{equation}
Which can be simplified into
\begin{equation}\label{eq:nextgen_adults_slimpify}
  \begin{aligned}
  p'  & = \hat{p}^2(T_B-T_A) + \hat{p}(1+T_A-T_B) \\
  \end{aligned}
  \end{equation}

The frequency of $A$ among parents follows a similar dynamic, but also includes the effect of natural selection, and is therefore
\begin{equation}\label{eq:nextgen_parents}
\begin{aligned}
\bar{w} \tilde{p}'
& = \hat{p}^2 (1+b-c) [\alpha + (1-\alpha)(\hat{p} + (1-\hat{p})(1-T_B))] \\
& + \hat{p}(1-\hat{p}) (1-c) [\alpha(1-T_B) + (1-\alpha)(\hat{p} + (1-\hat{p})(1-T_B))] \\
& + (1-\hat{p})\hat{p} (1+b) [\alpha T_A + (1-\alpha) \hat{p} T_A ] \\
& + (1-\hat{p})^2 [(1-\alpha) \hat{p} T_A],
\end{aligned}
\end{equation}
where $\bar{w} = \hat{p} w_A + (1-\hat{p}) w_B = 1 + \hat{p}(b-c)$ is the population mean fitness.
Eq~\ref{eq:nextgen_parents} can be slimpified to
\begin{equation}\label{eq:nextgen_parents_simplified}
\begin{aligned}
\bar{w} \tilde{p}'
& = \hat{p}^2 (1+b-c) [1-(1-\hat{p}(1-\alpha)T_B)] \\
& + \hat{p}(1-\hat{p}) (1-c) [\hat{p}(1-\alpha)T_B+1-T_B] \\
& + (1-\hat{p})\hat{p} (1+b) [\hat{p}(1-\alpha) + \alpha]T_A \\
& + (1-\hat{p})^2 [(1-\alpha) \hat{p}]T_A,
\end{aligned}
\end{equation}

>>>>

Horizontal cultural transmission occurs between peers. 
It may occur between social partners with probability $\alpha$, or between a random pair with probability $1-\alpha$ (see~\textbf{Fig.~\ref{fig:horizontal}}).
Horizontal transmission is not always successful, as one peer may reject the other's phenotype. The probability for successful transmission of phenotypes $A$ and $B$ are $T_A$ and $T_B$, respectively.\\
Table~\ref{table:1} contains the probability of first interactor to be A following interaction by interaction type.
Therefore, the frequency $p'$ of phenotype $A$ among adults in the next generation, after horizontal transmission, is 
\begin{equation}\label{eq:nextgen_adults}
\begin{aligned}
p'
& = \hat{p}^2 [\alpha + (1-\alpha)(\hat{p} + (1-\hat{p})(1-T_B))] \\
& + \hat{p}(1-\hat{p}) [\alpha(1-T_B) + (1-\alpha)(\hat{p} + (1-\hat{p})(1-T_B))] \\
& + (1-\hat{p})\hat{p} [\alpha T_A + (1-\alpha) \hat{p} T_A ] \\
& + (1-\hat{p})^2 [(1-\alpha) \hat{p} T_A]
\end{aligned}
\end{equation}
Which can be simplified into
\begin{equation}\label{eq:nextgen_adults_slimpify}
  \begin{aligned}
  p'  & = \hat{p}^2(T_B-T_A) + \hat{p}(1+T_A-T_B) \\
  \end{aligned}
  \end{equation}

The frequency of $A$ among parents follows a similar dynamic, but also includes the effect of natural selection, and is therefore
\begin{equation}\label{eq:nextgen_parents}
\begin{aligned}
\bar{w} \tilde{p}'
& = \hat{p}^2 (1+b-c) [\alpha + (1-\alpha)(\hat{p} + (1-\hat{p})(1-T_B))] \\
& + \hat{p}(1-\hat{p}) (1-c) [\alpha(1-T_B) + (1-\alpha)(\hat{p} + (1-\hat{p})(1-T_B))] \\
& + (1-\hat{p})\hat{p} (1+b) [\alpha T_A + (1-\alpha) \hat{p} T_A ] \\
& + (1-\hat{p})^2 [(1-\alpha) \hat{p} T_A],
\end{aligned}
\end{equation}
where fitness values are taken from table~\ref{table:1} and $\bar{w} = \hat{p} w_A + (1-\hat{p}) w_B = 1 + \hat{p}(b-c)$ is the population mean fitness.
Eq~\ref{eq:nextgen_parents} can be slimpified to
\begin{equation}\label{eq:nextgen_parents_simplified}
\begin{aligned}
\bar{w} \tilde{p}'
& = \hat{p}^2 (1+b-c) [1-(1-\hat{p}(1-\alpha)T_B)] \\
& + \hat{p}(1-\hat{p}) (1-c) [\hat{p}(1-\alpha)T_B+1-T_B] \\
& + (1-\hat{p})\hat{p} (1+b) [\hat{p}(1-\alpha) + \alpha]T_A \\
& + (1-\hat{p})^2 [(1-\alpha) \hat{p}]T_A,
\end{aligned}
\end{equation}

\begin{table}[]
\begin{tabular}{|c|c|c|c|c|}
\hline
\textbf{Interaction Type} & \textbf{Frequency}    & \textbf{Fitness}      & \multicolumn{2}{c|}{\textbf{\begin{tabular}[c]{@{}c@{}}Probability of first interactor\\  to be A following interaction\end{tabular}}}                                                                                                     \\ \hline
\multicolumn{1}{|l|}{}    & \multicolumn{1}{l|}{} & \multicolumn{1}{l|}{} & \textbf{\begin{tabular}[c]{@{}c@{}}Horizontal transmission\\  from partner, \\ probability $\alpha$\end{tabular}} & \textbf{\begin{tabular}[c]{@{}c@{}}Horizontal transmission \\ from population, \\ probability $1-\alpha$\end{tabular}} \\ \hline
$AxA$                     & $\hat{p}^2$           & $1+b-c$               & $1$                                                                                                               & $\hat{p}+(1-\hat{p})(1-T_B)$                                                                                           \\ \hline
$AxB$                     & $\hat{p}(1-\hat{p})$  & $1-c$                 & $1-T_B$                                                                                                           & $\hat{p}+(1-\hat{p})(1-T_B)$                                                                                           \\ \hline
$BxA$                     & $\hat{p}(1-\hat{p})$  & $1+b$                 & $T_A$                                                                                                             & $\hat{p}T_A$                                                                                                           \\ \hline
$BxB$                     & $(1-\hat{p})^2$       & $1$                   & $0$                                                                                                               & $\hat{p}T_A$                                                                                                           \\ \hline
\end{tabular}
\caption{Fitness and probabilityof first interactor to be A following interaction by interaction type}
\label{table:1}
\end{table}
